CUBE: Compact Particle-Mesh $N$-body Algorithm

### 2017-05-30 ms_caf/ms_caf.tex
haoran
add figure
cicpower (cross with x4v4)

\shortauthors{Author et al.} \shorttitle{New CubeP3M}
\begin{document}
\title{Compact Particle-Mesh $N$-body Algorithm}
\author{Authors}

\affil{Affiliations}

@@

\begin{abstract}
Cosmological large scale structure $N$-body simulations are computation-light but memory-heavy imbalanced problems in supercomputing. Traditional $N$-body simulation algorithms use at least 24-byte per particle in memory, where six 4-byte single precision real numbers keep track of each particle's phase space coordinate. This is usually the bottleneck of scaling-up of a $N$-body problem on modern supercomputers whose core/memory ratio is high. Here we present a parallel, memory-lite Particle-Mesh based $N$-body code, in which each particle can occupy as low as 6-byte memory. This is accomplished by storing only particles' relative location respect to a meshed grid, and their velocity relative to a predictable distribution function. The remaining information is given by the ordering of particles in memory space, and a complimentary density field, which is negligible in memory. Our numerical experiments show that this algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.

@@

\section{Introduction}
$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems \citep{1988csup.book.....H}. It is widely used in cosmology to model the formation and evolution of the large scale structure (LSS). With the fast development of parallel supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) $N$-body particles. To date the largest $N$-body simulation in application is the ``TianNu'' simulation run on the TianHe-2 supercomputer by cosmological simulation code {\tt CUBEP3M} \citep{2013MNRAS.436..540H}. It uses nearly $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age. However {\tt CUBEP3M} and TianNu did not fully use TianHe-2's computing power -- 87\% of the computing power on TianHe-2 are given by the Intel Xeon-Phi coprocessors while they occupy only 27\% of the total memory. Although we accomplished preliminary works on developing native and offload modes of {\tt CUBEP3M} on Xeon-Phi many integrated core (MIC) architecture, the memory problem is still the bottle neck of scaling up the $N$-body problem on the heterogeneous architectures. Lower the memory usage per particle will solve this problem, and will also help the $N$-body problem development on GPUs, which are even more massive parallelized and memory limited. In this paper we present a $N$-body algorithm which uses as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations. The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3. Discussions and conclusions are in \S 4.

@@

\begin{figure}[t]
\includegraphics[width=1.0\linewidth]{fig_tile.pdf}
\caption{Spacial decomposition.}
\label{fig.tile}
\end{figure}

@@

@@

\bibliographystyle{hapj}
\bibliography{../haoran_ref}

>>>>

\shortauthors{Yu et al.} \shorttitle{CUBE}
\begin{document}
\title{CUBE: Compact Particle-Mesh $N$-body Algorithm}
\author{
Hao-Ran Yu$^{1,2}$,
Ue-Li Pen$^{2,3,4,5}$,
Xin Wang$^{2}$
}

\affil{
$^1$Kavli Institute for Astronomy and Astrophysics, Peking University, Beijing 100871, China\\
$^2$Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5H 3H8 Canada\\
$^3$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, Toronto, ON M5S 3H4, Canada\\
$^4$Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, ON M5G 1Z8, Canada\\
$^5$Perimeter Institute for Theoretical Physics, Waterloo, ON N2L 2Y5, Canada
}

@@

\begin{abstract}
Cosmological large scale structure $N$-body simulations are computation-light but memory-heavy imbalanced problems in supercomputing. Traditional $N$-body simulation algorithms use at least 24-byte per particle in memory, where six 4-byte single precision real numbers keep track of each particle's phase space coordinate. This is usually the bottleneck of scaling-up of a $N$-body problem on modern supercomputers whose core/memory ratio is high. Here we present a parallel, memory-lite Particle-Mesh based $N$-body code, in which each particle can occupy as low as 6-byte memory. This is accomplished by storing only particles' relative location respect to a mesh-grid, and their velocity relative to a predictable distribution function. The remaining information is given by the ordering of particles, and a complimentary density field, which is negligible in memory space. Our numerical experiments show that this algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.

@@

\begin{figure}[t]
\centering
  \includegraphics[width=0.95\linewidth]{ms_caf.pdf}
\end{figure}
\section{Introduction}
$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems \citep{1988csup.book.....H}. It is widely used in cosmology to model the formation and evolution of the large scale structure (LSS). With the fast development of parallel supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) $N$-body particles. To date the largest $N$-body simulation in application is the ``TianNu'' simulation run on the TianHe-2 supercomputer by cosmological simulation code {\tt CUBEP3M} \citep{2013MNRAS.436..540H}. It uses nearly $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age. However {\tt CUBEP3M} and TianNu did not fully use TianHe-2's computing power -- 87\% of the computing power on TianHe-2 are given by the Intel Xeon-Phi coprocessors while they occupy only 27\% of the total memory. Although we accomplished preliminary works on developing native and offload modes of {\tt CUBEP3M} on Xeon-Phi many integrated core (MIC) architecture, the memory problem is still the bottle neck of scaling up the $N$-body problem on the heterogeneous architectures. Lower the memory usage per particle will solve this problem, and will also help the $N$-body problem development on GPUs, which are even more massive parallelized and memory limited. In this paper we present a $N$-body algorithm which uses as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations. The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3. Discussions and conclusions are in \S 4.

@@

\begin{figure}[t]
\centering
  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
 \caption{Spacial decomposition.}
\label{fig.tile}
\end{figure}

@@

\begin{figure}[t]
\centering
  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
 \caption{Displacement.}
\label{fig.dsp}
\end{figure}

@@

\bibliographystyle{hapj}
\bibliography{haoran_ref}
